---
title: "ADS 2017 Presentation Analyses"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    toc_depth: 4
    code_folding: hide
    theme: yeti
---

```{r setup}
library(tidyverse)
library(rms)

## Color scheme to match Keynote presentation
gray.colour <- "#CDCDCD"
back.colour <- "#222222"
kn.colours <- c("blue" = "#38A8DA", "yellow" = "#FECB55", "green" = "#89BF60")

```

Notes:

- All graphics are optimized for a presentation, not this Markdown document. If they look weird in this document, that's why.
- All random samples are done without setting seeds and will look different every time the program is run. So if figures look different in this document than presentation, that's why.
- Work on one of these projects is ongoing; therefore, only code (not results) is publicly provided.

# TBI Cohort Descriptives

```{r tbi_prep}
load('Data/tbi_datasets.Rdata')

## Cohort in TBI vs mortality is restricted to patients with blunt injury, >=1 day in the ICU, and
## at least one day of EMR data
tbi.icupts <- subset(tbi.oneobs, pt.injury == 'Blunt' & icu.los >= 1 & mrn %in% tbi.daily$mrn)$mrn

tbi.oneobs <- subset(tbi.oneobs, mrn %in% tbi.icupts)
tbi.daily <- subset(tbi.daily, mrn %in% tbi.icupts)

## How many *days* were missing mental status assessments?
tbi.daily$had.asmt <- !is.na(tbi.daily$mental.status)

n.days.miss <- sum(tbi.daily$had.asmt)
pct.days.miss <- round(mean(tbi.daily$had.asmt)*100)

## How many *patients* were completely, partially, or not missing data?
tbi.oneobs <- tbi.oneobs %>%
  mutate(partial.asmt = days.mental > 0 & days.mental < n.recs,
         no.asmt = days.mental == 0,
         comp.asmt = days.mental == n.recs,
         had.asmt = partial.asmt | comp.asmt,
         asmt.data = factor(ifelse(no.asmt, 1,
                            ifelse(comp.asmt, 3, 2)),
                            levels = 1:3,
                            labels = c('All Missing',
                                       'Some Missing',
                                       'Complete Data')))

n.comp.miss <- sum(tbi.oneobs$no.asmt)
pct.comp.miss <- mean(tbi.oneobs$no.asmt)

n.partial.miss <- sum(tbi.oneobs$partial.asmt)
pct.partial.miss <- mean(tbi.oneobs$partial.asmt)

n.none.miss <- sum(tbi.oneobs$comp.asmt)
pct.none.miss <- mean(tbi.oneobs$comp.asmt)

n.atleast1.miss <- n.partial.miss + n.comp.miss
pct.atleast1.miss <- n.atleast1.miss / length(tbi.icupts)

## Create table to describe patient-level missingness
pt.missing <- tribble(
  ~ group, ~ pct, ~ n,
  "Total N", NA, length(tbi.icupts),
  "Complete assessment data", pct.none.miss, n.none.miss,
  "Missing at least one assessment", pct.atleast1.miss, n.atleast1.miss,
  "Partial assessment data", pct.partial.miss, n.partial.miss,
  "No assessment data", pct.comp.miss, n.comp.miss
) %>%
  mutate(pct = round(pct, 2))

pt.missing

```

We have `r length(tbi.icupts)` who were in the ICU with electronic medical record data available for at least one day; these patients had `r nrow(tbi.daily)` total patient-days. Of these patient-days, `r n.days.miss` (`r pct.days.miss`%) were missing mental status assessment.

## How Were Patients with Mental Status Assessments Different from Patients Without?
```{r tbi_describe_pts}
html(summaryM(age + gender + race + insurance.code + iss + cpr.yn + pt.marshall + pt.cerebral +
                pt.epidural + vent.days + disposition.coded + fim.total + icu.los + los + hosp.death +
                base.motor + base.pupil.react + n.recs ~ asmt.data,
              data = tbi.oneobs),
     digits = 2)

```

We see differences in age (patients without assessments are older); insurance (more likely to have public insurance); injury severity (have lower ISS); Marshall Class (more likely to be Marshall Class II vs I or III); cerebral SAH (less likely to have this); time on the vent (much less likely to be ventilated); discharge disposition (more likely to be discharged home, less likely to die in hospital/go to rehab); ICU and hospital LOS (shorter); max motor response (higher); pupil reactivity (less likely to have both fixed). Let's plot some of these to show our friends at ADS.

```{r tbi_pt_plots}
## -- Function to create consistent boxplots -------------------------------------------------------
tbi.boxplots <- function(yvar, ylims = NULL, df){
  p <- ggplot(data = df, aes_string(x = 'had.asmt', y = yvar)) +
    geom_point(aes(colour = had.asmt), alpha = 0.4, position = position_jitter()) +
    geom_boxplot(aes(fill = had.asmt), alpha = 0.5, colour = gray.colour, outlier.shape = NA) +
    scale_fill_manual(values = as.character(kn.colours[c('green', 'blue')]), guide = FALSE) +
    scale_colour_manual(values = as.character(kn.colours[c('green', 'blue')]), guide = FALSE) +
    scale_x_discrete(labels = c('No assessment', '>=1'), name = NULL) +
    scale_y_continuous(name = NULL) +
    theme_minimal() +
    theme(plot.background = element_rect(fill = 'transparent', colour = NA),
          axis.text = element_text(colour = gray.colour, size = 10),
          panel.grid.major = element_line(colour = gray.colour),
          panel.grid.minor = element_line(colour = gray.colour))

  if(!is.null(ylims)){
    p <- p +
      scale_y_continuous(name = NULL, limits = ylims)
  } else{
    p <- p +
      scale_y_continuous(name = NULL)
  }

  p
}

tbi.age.box <- tbi.boxplots('age', df = tbi.oneobs)
tbi.iss.box <- tbi.boxplots('iss', df = tbi.oneobs)
tbi.vent.box <- tbi.boxplots('vent.days', ylims = c(0, 40), df = tbi.oneobs)
tbi.los.box <- tbi.boxplots('los', ylims = c(0, 100), df = tbi.oneobs)

## -- Function to save each boxplot as a PDF to insert in Keynote ----------------------------------
save.tbi.boxplots <- function(p, fname){
  ggsave(filename = paste0(fname, '.pdf'),
         plot = p, device = 'pdf', path = 'ADSFigures', width = 3, height = 3)
}

save.tbi.boxplots(tbi.age.box, 'tbi_age_box')
save.tbi.boxplots(tbi.iss.box, 'tbi_iss_box')
save.tbi.boxplots(tbi.vent.box, 'tbi_vent_box')
save.tbi.boxplots(tbi.los.box, 'tbi_los_box')

```

## How Were Patient-Days with Missing Status Different from Those Without?

`FALSE` = no mental status available; `TRUE` = patient had at least one mental status assessment that day.

```{r tbi_describe_days}
html(summaryM(max.motor.imp + pupil.react.imp + min.hemoglobin.imp + max.icp.imp + max.icp.dich +
                sofa.nanormal.imp + sofa.namissing.imp + tot.benzo + tot.opioid + tot.propofol +
                tot.dex + tot.antipsyc + tot.betablock + tot.clonid + units.cryo + units.plasma +
                units.platelets + units.prbc ~ !is.na(mental.status),
              data = tbi.daily),
     digits = 2)

```

We see differences in maximum motor response, SOFA, and benzodiazepine and opioid doses.

```{r tbi_daily_box}
## -- Boxplots for daily data ----------------------------------------------------------------------
tbi.motor.box <- tbi.boxplots('max.motor.imp', df = tbi.daily)
tbi.sofa.box <- tbi.boxplots('sofa.nanormal.imp', df = tbi.daily)
tbi.benzo.box <- tbi.boxplots('tot.benzo', df = tbi.daily)
tbi.opioid.box <- tbi.boxplots('tot.opioid', df = tbi.daily)

save.tbi.boxplots(tbi.motor.box, 'tbi_motor_box')
save.tbi.boxplots(tbi.sofa.box, 'tbi_sofa_box')
save.tbi.boxplots(tbi.benzo.box, 'tbi_benzo_box')
save.tbi.boxplots(tbi.opioid.box, 'tbi_opioid_box')

```

## Differences in Delirium Duration Depending on Different Assumptions
One of our main exposures for long-term outcomes is duration of delirium in the index hospitalization. What does that outcome look like if we...

```{r deldur}
## -- Demonstrate how delirium durations change based on how we treat missing days -----------------
tbi.delassume <- tbi.daily %>%
  group_by(mrn) %>%
  summarise(days.available = sum(!is.na(mental.status)),
            ## Ignore missing; count up all available delirious assessments
            days.del.ignore = ifelse(days.available == 0, NA,
                                     sum(mental.status == 'Delirious', na.rm = TRUE)),
            ## Missing = normal
            days.del.normal = sum(mental.status == 'Delirious', na.rm = TRUE),
            ## Missing = delirious
            days.del.del = sum(is.na(mental.status) | mental.status == 'Delirious')) %>%
  mutate(diff.per.pt = days.del.del - days.del.normal)

## Get descriptive stats for each assumption
tbi.assumed.ignore <- sprintf("%s (%s, %s); N = %s",
                              median(tbi.delassume$days.del.ignore, na.rm = TRUE),
                              quantile(tbi.delassume$days.del.ignore, probs = 0.25, na.rm = TRUE),
                              quantile(tbi.delassume$days.del.ignore, probs = 0.75, na.rm = TRUE),
                              sum(!is.na(tbi.delassume$days.del.ignore)))

tbi.assumed.normal <- sprintf("%s (%s, %s); N = %s",
                              median(tbi.delassume$days.del.normal, na.rm = TRUE),
                              quantile(tbi.delassume$days.del.normal, probs = 0.25, na.rm = TRUE),
                              quantile(tbi.delassume$days.del.normal, probs = 0.75, na.rm = TRUE),
                              sum(!is.na(tbi.delassume$days.del.normal)))

tbi.assumed.del <- sprintf("%s (%s, %s); N = %s",
                           median(tbi.delassume$days.del.del, na.rm = TRUE),
                           quantile(tbi.delassume$days.del.del, probs = 0.25, na.rm = TRUE),
                           quantile(tbi.delassume$days.del.del, probs = 0.75, na.rm = TRUE),
                           sum(!is.na(tbi.delassume$days.del.del)))

## Difference between delirious & normal assumptions
tbi.assumed.diff <- sprintf("%s (%s, %s)",
                            median(tbi.delassume$diff.per.pt, na.rm = TRUE),
                            quantile(tbi.delassume$diff.per.pt, probs = 0.25, na.rm = TRUE),
                            quantile(tbi.delassume$diff.per.pt, probs = 0.75, na.rm = TRUE))

```

- Ignore missingness; just sum up days of delirium per patient: `r tbi.assumed.ignore`
- Assume missing = normal: `r tbi.assumed.normal`
- Assume missing = delirium: `r tbi.assumed.del`
- Difference per patient of `r tbi.assumed.diff` days

```{r deldur_plot}
tbi.delassume.plot <- tbi.delassume %>%
  gather(key = assumption, value = days, days.del.ignore:days.del.del) %>%
  mutate(assumption.f = factor(ifelse(assumption == 'days.del.ignore', 1,
                               ifelse(assumption == 'days.del.normal', 2,
                               ifelse(assumption == 'days.del.del', 3, NA))),
                               levels = 1:3,
                               labels = c(paste0('Missing days ignored\n', tbi.assumed.ignore),
                                          paste0('Missing = Normal\n', tbi.assumed.normal),
                                          paste0('Missing = Delirious\n', tbi.assumed.del))),
         dummy = 1)
         
## -- Create boxplots for distributions of delirium durations with different assumptions -----------
ggplot(data = subset(tbi.delassume.plot, assumption != 'days.del.normal'),
         ## decided not to include missing = normal in plot; not that different from ignoring
       aes(x = dummy, y = days)) +
  facet_wrap(~ assumption.f, nrow = 3) +
  geom_point(aes(colour = assumption.f), alpha = 0.4, position = position_jitter()) +
  geom_boxplot(aes(fill = assumption.f),
               alpha = 0.5, colour = gray.colour, size = 1, outlier.shape = NA) +
  scale_y_continuous(limits = c(0, 30)) +
  scale_colour_manual(values = as.character(kn.colours[c(3, 1)])) +
  scale_fill_manual(values = as.character(kn.colours[c(3, 1)])) +
  coord_flip() +
  theme_minimal() +
  theme(plot.background = element_rect(fill = 'transparent', colour = NA),
        axis.text.y = element_blank(),
        axis.text.x = element_text(colour = 'grey50', size = 12, face = 'bold'),
        axis.title = element_blank(),
        panel.grid.major = element_line(colour = 'grey50'),
        panel.grid.minor = element_line(colour = 'grey50'),
        strip.text = element_text(colour = gray.colour, size = 18, face = 'bold'),
        legend.position = 'none')

ggsave(filename = 'tbi_delassume.pdf', device = 'pdf', path = 'ADSFigures', width = 5, height = 7)

```

# BRAIN-ICU Modifications

```{r brain_prep}
load('Data/braindata.Rdata')

## brain.daily includes 30 obs/patient, including records for after discharge/withdrawal/death;
## create indicator for whether patient was in hospital + still in study on each day
brain.daily$hospital.day <- brain.daily$status.today.imp %in% c('Normal', 'Delirious', 'Comatose')

n.days.miss <- sum(is.na(subset(brain.daily, hospital.day)$mental.stat))
pct.days.miss <- mean(is.na(subset(brain.daily, hospital.day)$mental.stat))

n.pts.miss <- sum(brain.oneobs$id %in% subset(brain.daily, hospital.day & is.na(mental.stat))$id)
pct.pts.miss <- mean(brain.oneobs$id %in% subset(brain.daily, hospital.day & is.na(mental.stat))$id)

## How many patient-days were missing per patient missing *any*?
brain.missdays <- brain.daily %>%
  filter(hospital.day & is.na(mental.stat)) %>%
  group_by(id) %>%
  summarise(days.miss = n())

brain.missdays.stats <- sprintf("%s (%s, %s)",
                                median(brain.missdays$days.miss),
                                quantile(brain.missdays$days.miss, probs = 0.25),
                                quantile(brain.missdays$days.miss, probs = 0.75))

```

In BRAIN-ICU, we had `r round(pct.days.miss*100, 2)`% of patient-days missing; this missingness was spread among `r n.pts.miss` patients (`r round(pct.pts.miss*100)`% of our total cohort). Among patients missing any data, the median (IQR) days missing was `r brain.missdays.stats`.

We want to see how our results change if we assume different amounts of delirium days were 1) missing completely at random and 2) missing at random.

```{r brain_delmissing}
has.rbans <- subset(brain.fu, fu.period == '3 Month' & !is.na(rbans.global.score))$id

## Cube root drug doses
brain.oneobs <- brain.oneobs %>%
  filter(id %in% has.rbans) %>%
  mutate(mean.benz.cube = mean.benz.icu^(1/3),
         mean.op.new.cube = mean.op.new.icu^(1/3),
         mean.prop.cube = mean.prop.icu^(1/3),
         mean.dex.cube = mean.dex.icu^(1/3),
         mean.hal.cube = mean.hal.icu^(1/3))

## Set up variables for 10/25/50% missing delirium days, both MCAR and MAR
brain.daily$mental.stat.mcar.10 <- brain.daily$mental.stat.mcar.25 <-
  brain.daily$mental.stat.mcar.50 <- brain.daily$mental.stat.mar.10 <-
  brain.daily$mental.stat.mar.25 <- brain.daily$mental.stat.mar.50 <- brain.daily$mental.stat.imp

## We only need in-hospital days, not entire study period
brain.hospital <- brain.daily %>%
  filter(hospital.day & id %in% has.rbans) %>%
  mutate(mean.prop.cube = mean.prop.icu^(1/3),
         mean.benz.cube = mean.benz.icu^(1/3),
         mean.op.new.cube = mean.op.new.icu^(1/3))

## How many hospital days were there?
hosp.days <- sum(brain.hospital$hospital.day)

## Easy ones: MCAR
brain.hospital$mental.stat.mcar.10[sample(1:nrow(brain.hospital), size = round(hosp.days*0.1))] <-
  NA
brain.hospital$mental.stat.mcar.25[sample(1:nrow(brain.hospital), size = round(hosp.days*0.25))] <-
  NA
brain.hospital$mental.stat.mcar.50[sample(1:nrow(brain.hospital), size = round(hosp.days*0.5))] <-
  NA

## Harder: MAR - predict daily missingness based on age, SOI, opioids, propofol
brain.hospital <- brain.hospital %>%
  left_join(dplyr::select(brain.oneobs, id, age.enroll), by = 'id')

## Ended up just predicting with propofol
mod.missing.original <- lrm(is.na(mental.stat) ~ rcs(mean.prop.cube, 3),
                            data = brain.hospital)

## (If we were doing this for a full analysis, we'd need to account for clustering by patient)

## Calculate P(missing) for each patient-day
brain.hospital$missing.prob <-
  predict(mod.missing.original, newdata = brain.hospital, type = 'lp') %>% plogis

## Calculate quantiles of missingness probabilities
brain.mar.qs <- quantile(brain.hospital$missing.prob, probs = c(0.9, 0.75, 0.5))

## For each % missing, set all days with P(missing) > quantile to NA
brain.hospital$mental.stat.mar.10[brain.hospital$missing.prob > brain.mar.qs['90%']] <- NA
brain.hospital$mental.stat.mar.25[brain.hospital$missing.prob > brain.mar.qs['75%']] <- NA
brain.hospital$mental.stat.mar.50[brain.hospital$missing.prob > brain.mar.qs['50%']] <- NA

## Get delirium durations for each value of missingness so we can run models with each
brain.deldur.missing <- brain.hospital %>%
  group_by(id) %>%
  summarise(del.mcar.10 = sum(mental.stat.mcar.10 == 'Delirious', na.rm = TRUE),
            coma.mcar.10 = sum(mental.stat.mcar.10 == 'Comatose', na.rm = TRUE),
            del.mcar.25 = sum(mental.stat.mcar.25 == 'Delirious', na.rm = TRUE),
            coma.mcar.25 = sum(mental.stat.mcar.25 == 'Comatose', na.rm = TRUE),
            del.mcar.50 = sum(mental.stat.mcar.50 == 'Delirious', na.rm = TRUE),
            coma.mcar.50 = sum(mental.stat.mcar.50 == 'Comatose', na.rm = TRUE),
            del.mar.10 = sum(mental.stat.mar.10 == 'Delirious', na.rm = TRUE),
            coma.mar.10 = sum(mental.stat.mar.10 == 'Comatose', na.rm = TRUE),
            del.mar.25 = sum(mental.stat.mar.25 == 'Delirious', na.rm = TRUE),
            coma.mar.25 = sum(mental.stat.mar.25 == 'Comatose', na.rm = TRUE),
            del.mar.50 = sum(mental.stat.mar.50 == 'Delirious', na.rm = TRUE),
            coma.mar.50 = sum(mental.stat.mar.50 == 'Comatose', na.rm = TRUE)) %>%
  ## Join baseline variables so we can mimic original BRAIN model
  right_join(dplyr::select(brain.oneobs, id, age.enroll, charlson.score, edu, iqcode.score.e,
                           stroke.risk, e4.type, mean.modsofa.icu.old, icudays.sevseptic.s,
                           ints.sat.l90.s, mean.benz.cube, mean.op.new.cube, mean.prop.cube,
                           mean.hal.cube, mean.dex.cube, del.s.imp, coma.s.imp),
             by = 'id') %>%
  left_join(dplyr::select(filter(brain.fu, fu.period == '3 Month'), id, rbans.global.score))

## Function to run a model given a type (MCAR, MAR) and % missing (10, 25, 50)
run.miss.mod <- function(mtype = c('mcar', 'mar'),
                         mpct = c('0', '10', '25', '50')){
  
  mtype <- match.arg(mtype)
  mpct <- match.arg(mpct)
  
  if(mpct == '0'){
    delvar <- 'del.s.imp'
    comavar <- 'coma.s.imp'
  } else{
    delvar <- paste('del', mtype, mpct, sep = '.')
    comavar <- paste('coma', mtype, mpct, sep = '.')
  }
  
  modform <- as.formula(sprintf("rbans.global.score ~ rcs(age.enroll, 4) + rcs(charlson.score, 4) + rcs(edu, 4) + rcs(iqcode.score.e, 4) + rcs(stroke.risk, 4) + e4.type + rcs(mean.modsofa.icu.old, 4) + rcs(icudays.sevseptic.s, 4) + rcs(ints.sat.l90.s, 4) + rcs(%s, 4)*rcs(%s, 4) + rcs(mean.benz.cube, 4) + rcs(mean.op.new.cube, 4) + rcs(mean.prop.cube, 4) + mean.dex.cube + mean.hal.cube",
                        delvar, comavar))
  
  list(ols(modform, data = brain.deldur.missing), delvar)
}

## List of original + MCAR + MAR models
brain.missing.mods <- map2(list('mcar', 'mcar', 'mcar', 'mcar', 'mar', 'mar', 'mar'),
                           list('0', '10', '25', '50', '10', '25', '50'),
                           .f = run.miss.mod)

## -- Function to take one element of brain.missing.mods and calculate predicted RBANS for each ----
## -- delirium duration ----------------------------------------------------------------------------

## Gotta set that datadist to use rms::Predict()
dd <- datadist(brain.deldur.missing); options(datadist = 'dd')

calc.predvals <- function(listobj){
  mod <- listobj[[1]]
  delvar <- listobj[[2]]
  
  ## Call Predict() using a given model and delirium duration variable
  predvals <- eval(parse(text = sprintf("Predict(mod, %s = NA)", delvar))) %>%
    as.data.frame() %>%
    ## Create variables for types/amounts of missing data
    mutate(delvar = ifelse(delvar == 'del.s.imp', 'original', gsub('^del\\.', '', delvar))) %>%
    separate(delvar, into = c('mtype', 'mpct'), remove = FALSE) %>%
    ## Format missingness variables for facet labels
    mutate(mtype = ifelse(mtype == 'mcar', 'Missing Completely\nat Random (MCAR)',
                   ifelse(mtype == 'mar', 'Missingness Related to\nCovariates (MAR)', NA)),
           mpct = ifelse(is.na(mtype), NA, paste0(mpct, "% Missing")))
  
  ## Rename delirium exposures to "delirium" in each dataset to make plotting function easier
  names(predvals) <- gsub(delvar, 'delirium', names(predvals))
  
  predvals
}

## -- Use above function to get a big data frame with all our predicted values ---------------------
brain.missing.predvals <- map_df(brain.missing.mods, .f = calc.predvals)


## -- Figure for original + modified model results -------------------------------------------------
## Set values for X axis limits
del.limits <- c(0, 25)

## Data management to add original data to each facet: need faceting values (type/amount)
brain.mod.org.faceted <-  map2(list('Missing Completely\nat Random (MCAR)',
                                    'Missing Completely\nat Random (MCAR)',
                                    'Missing Completely\nat Random (MCAR)',
                                    'Missingness Related to\nCovariates (MAR)',
                                    'Missingness Related to\nCovariates (MAR)',
                                    'Missingness Related to\nCovariates (MAR)'),
                               list('10% Missing', '25% Missing', '50% Missing',
                                    '10% Missing', '25% Missing', '50% Missing'),
                               .f = ~ subset(brain.missing.predvals, delvar == 'original') %>%
                                 mutate(mtype = .x, mpct = .y)) %>%
  bind_rows()

## Create main figure
brain.mod.fig <- ggplot(data = subset(brain.missing.predvals, delvar != 'original'),
                        aes(x = delirium, y = yhat)) +
  facet_grid(mtype ~ mpct) +
  ## Modified results (CI + point estimate)
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = mtype), alpha = 0.5) +
  geom_line(aes(colour = mtype), size = 2) +
  ## Original results (CI + point estimate)
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.3, fill = gray.colour,
              data = brain.mod.org.faceted) +
  geom_line(data = brain.mod.org.faceted, alpha = 0.4, colour = 'gray30', size = 2) +
  scale_colour_manual(values = as.character(kn.colours[c(1, 3)])) +
  scale_fill_manual(values = as.character(kn.colours[c(1, 3)])) +
  scale_x_continuous(name = 'Days of Delirium', limits = del.limits) +
  scale_y_continuous(name = 'Adjusted RBANS Global Score', limits = c(35, 100)) +
  labs(caption = '\nGrayscale = original study results; color = modified as noted') +
  theme_minimal() +
  theme(plot.background = element_rect(fill = 'transparent', colour = NA),
        axis.text = element_text(colour = gray.colour, size = 14),
        axis.title = element_text(colour = gray.colour, size = 14),
        plot.title = element_text(colour = gray.colour, size = 20),
        plot.caption = element_text(colour = gray.colour, face = 'italic', size = 16),
        strip.text = element_text(colour = gray.colour, size = 16),
        panel.spacing = unit(0.5, 'inches'),
        panel.grid.major = element_line(colour = 'gray30'),
        panel.grid.minor = element_line(colour = 'gray30'),
        panel.border = element_rect(fill = NA, colour = 'gray30'),
        legend.position = 'none')

brain.mod.fig

ggsave(filename = 'brain_modifiedmods.pdf',
       device = 'pdf', width = 10, height = 7, path = 'ADSFigures')

```

# Visuals for Types of Missingness

Create dotplot of 1000 observations with a random X value (eg, age), and sorted/colored by three groups (unmeasured confounder).

- MCAR: make missing completely at random
- MAR: make missingness at higher rate with higher X
- MNAR: make missing depending on unmeasured confounder

```{r random_types}
## Base data set: variables for X, Y, category (eg, study site)
random_nums <- data.frame(x = flatten_int(map(1:30, ~rep(., times = 30))),
                          y = rep(1:30, 30),
                          category = c(rep('A', 300), rep('B', 300), rep('C', 300)),
                          stringsAsFactors = FALSE)

## Set value for point size (optimized for Keynote, not RMarkdown!)
ptsize <- 7

## Setup plot: all X/Y values by category
base_plot <- ggplot(data = random_nums, aes(x = x, y = y)) +
  geom_point(aes(colour = category), size = ptsize) +
  scale_colour_manual(values = as.character(kn.colours)) +
  theme_minimal() +
  theme(plot.background = element_rect(fill = 'transparent', colour = NA),
        legend.position = 'none',
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank())

## MCAR: Set 1/3 values to NA at random
mcar_nums <- random_nums
mcar_nums$category[sample(1:nrow(mcar_nums), size = 300)] <- 'Missing'

mcar_plot <- ggplot(data = mcar_nums, aes(x = x, y = y)) +
  geom_point(aes(colour = category), size = ptsize) +
  scale_colour_manual(values = c(as.character(kn.colours), back.colour)) +
  theme_minimal() +
  theme(plot.background = element_rect(fill = 'transparent', colour = NA),
        legend.position = 'none',
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank())

mcar_plot

## MAR: missingness varies by category
mar_is_missing <- c(as.logical(rbinom(n = 300, size = 1, prob = 0.75)),
                    as.logical(rbinom(n = 300, size = 1, prob = 0.4)),
                    as.logical(rbinom(n = 300, size = 1, prob = 0.1)))

mar_nums <- cbind(random_nums, mar_is_missing) %>%
  mutate(category = ifelse(mar_is_missing, 'Missing', category))

mar_plot <- ggplot(data = mar_nums, aes(x = x, y = y)) +
  geom_point(aes(colour = category), size = ptsize) +
  scale_colour_manual(values = c(as.character(kn.colours), back.colour)) +
  theme_minimal() +
  theme(plot.background = element_rect(fill = 'transparent', colour = NA),
        legend.position = 'none',
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank())

mar_plot

## MNAR: missingness varies with unmeasured confounder (x axis)
create_mnar_indicator <- function(mdenom){
  tmp <- rep(FALSE, 30)
  tmp[sample(1:30, size = ((1 / mdenom)*30))] <- TRUE
  tmp
}

mnar_is_missing <- map(rep(1:10, 3), create_mnar_indicator) %>%
  flatten_lgl()

mnar_nums <- cbind(random_nums, mnar_is_missing) %>%
  mutate(category = ifelse(mnar_is_missing, 'Missing', category))

## Add shuffled X value to demonstrate what MNAR data would look like to our eyeballs
rand_order <- sample(1:900, size = 900, replace = FALSE)
mnar_nums$xrand <- mnar_nums$x[rand_order]
mnar_nums$yrand <- mnar_nums$y[rand_order]

## Plot if we knew X
mnar_plot <- ggplot(data = mnar_nums, aes(x = x, y = y)) +
  geom_point(aes(colour = category), size = ptsize) +
  scale_colour_manual(values = c(as.character(kn.colours), back.colour)) +
  theme_minimal() +
  theme(plot.background = element_rect(fill = 'transparent', colour = NA),
        panel.background = element_rect(fill = 'transparent', colour = gray.colour),
        legend.position = 'none',
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank())

mnar_plot

## Plot if we don't know X
mnar_plot_xunknown <-  ggplot(data = mnar_nums, aes(x = xrand, y = yrand)) +
  geom_point(aes(colour = category), size = ptsize) +
  scale_colour_manual(values = c(as.character(kn.colours), back.colour)) +
  theme_minimal() +
  theme(plot.background = element_rect(fill = 'transparent', colour = NA),
        panel.background = element_rect(fill = 'transparent', colour = gray.colour),
        legend.position = 'none',
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank())

mnar_plot_xunknown

ggsave(filename = 'base.pdf', base_plot, path = 'ADSFigures', device = 'pdf', width = 10, height = 8)
ggsave(filename = 'mcar.pdf', mcar_plot, path = 'ADSFigures', device = 'pdf', width = 10, height = 8)
ggsave(filename = 'mar.pdf', mar_plot, path = 'ADSFigures', device = 'pdf', width = 10, height = 8)
ggsave(filename = 'mnar.pdf', mnar_plot, path = 'ADSFigures', device = 'pdf', width = 10, height = 8)
ggsave(filename = 'mnar_xunknown.pdf', mnar_plot_xunknown,
       path = 'ADSFigures', device = 'pdf', width = 10, height = 8)

```
